Spin-orbit-coupling-induced squeezing in the optical lattice clock

### 2018-07-10 soc_squeezing/soc_squeezing_notes.tex
fixed misnamed figure
tweaked example parameters for OAT and TAT

\begin{table}[h]
  \centering
  \caption{Example parameters for spin squeezing via one-axis
    twisting.}
  \label{tab:parameters}
  \begin{tabular}{|l|c|l|}
    \hline
    Parameter & Symbol & Value \\ \hline\hline
    Primary lattice depth & $V_0$ & 4 $E_R$ \\
    Transverse lattice depths & $V_T$ & 80 $E_R$ \\
    Lattice sites & $L$ & 100 \\
    Atom number & $N$ & 50 \\
    SOC strength & $\phi$ & $\pi/50$ \\ \hline\hline
    Tunneling rate & $J_0$ & $\approx300\times2\pi$ Hz \\
    On-site interaction strength & $U$ & $\approx1.4\times2\pi$ kHz \\
    Collective interaction strength & $\xi$
    & $\approx14\times2\pi$ Hz \\
    Squeezing strength & $\chi$ & $\approx21\times2\pi$ mHz \\
    Optimal squeezing time & $t_{\t{opt}}^{\t{OAT}}$
    & $\approx0.66$ seconds \\ \hline
  \end{tabular}
\end{table}

@@

Rather than pulsing the clock laser, we can apply a continuous drive
to realize the Hamiltonian
\begin{align}
  H_{\t{drive}} = H_{\t{free}} + H_{\t{clock}}
  = -\xi\v S\c\v S - \bar h S_z - \Omega S_x - \sum_n b_n s_n^z,
\end{align}
where the drive strength $\Omega$ may generally depend on time.  As
shown in Appendix \ref{sec:effective_squeezing_drive}, the addition of
a drive with $\abs{\Omega}\ll N\abs{\xi}$ does not modify the
squeezing behavior of the spin-orbit coupling terms in
$H_{\t{drive}}$, whose effective action on the $S=N/2$ manifold is
captured by
\begin{align}
  H_{\t{drive,eff}} = -\bar h S_z - \Omega S_x + \chi S_z^2.
  \label{eq:H_drive_eff}
\end{align}
To follow the continuous-drive two-axis twisting prescription in
ref.~\cite{huang2015twoaxis}, we now
\begin{enumerate*}[label=(\roman*)]
\item blue-detune the clock laser by $\bar h$ and shift into an
  appropriate rotating frame to eliminate the $\sim S_z$ term in
  \eqref{eq:H_drive_eff},
\item modulate the drive as $\Omega\p{t}=\beta\omega\cos\p{\omega t}$
  with $\omega\gg N\chi$, and
\item move into the rotating frame of $-\Omega\p{t}S_x$.
\end{enumerate*}
After a secular approximation which relies on $\omega\gg N\chi$, this
procedure results in the effective Hamiltonian
\begin{align}
  H_{\t{drive,eff}}^{\t{mod}}
  = \f{\chi}{2} \p{\sp{\J_0\p{2\beta}+1} S_z^2
    - \sp{\J_0\p{2\beta}-1} S_y^2},
\end{align}
where $\J_0$ is the zero-order Bessel function of the first kind.
Choosing a modulation index $\beta$ such that $\J_0\p{2\beta}=1/3$ or
$\J_0\p{2\beta}=-1/3$ then respectively results in the two-axis
twisting Hamiltonians
\begin{align}
  H_{\t{TAT}}^{z,x}
  = \f{\chi}{3} \p{2 S_z^2 + S_y^2}
  \simeq \f{\chi}{3} \p{S_z^2 - S_x^2},
  &&
  H_{\t{TAT}}^{y,x}
  = \f{\chi}{3} \p{S_z^2 + 2 S_y^2}
  \simeq \f{\chi}{3}\p{S_y^2 - S_x^2},
  \label{eq:H_TAT_drive}
\end{align}
where $\simeq$ denotes equality up to a global energy shift. Given the
set of parameters in Table \ref{tab:parameters}, the two-axis twisting
Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with
$\omega=20\times2\pi$ Hz and $\beta=0.906$.  The optimal squeezing
time for this protocol is $t_{\t{opt}}^{\t{TAT}}\approx0.98$ seconds.
A comparison of squeezing through $H_{\t{OAT}}$ and
$H_{\t{TAT}}^{z,x}$ via the parameters in Table \ref{tab:parameters}
is provided in Figure \ref{fig:squeezing_comparison}.

\begin{figure}[h]
  \centering
  \includegraphics{squeezing_comparison.pdf}
  \caption{Comparison of squeezing through $H_{\t{OAT}}$ and
    $H_{\t{TAT}}^{z,x}$ via the parameters in Table
    \ref{tab:parameters} and the continuous two-axis twisting protocol
    with $\omega=20\times2\pi$ Hz and $\beta=0.906$.}
  \label{fig:parameters}
\end{figure}

>>>>

\begin{table}[h]
  \centering
  \caption{Example parameters for spin squeezing via one-axis
    twisting.}
  \label{tab:parameters}
  \begin{tabular}{|l|c|l|}
    \hline
    Parameter & Symbol & Value \\ \hline\hline
    Primary lattice depth & $V_0$ & 4 $E_R$ \\
    Transverse lattice depths & $V_T$ & 60 $E_R$ \\
    Lattice sites & $L$ & 100 \\
    Atom number & $N$ & 50 \\
    SOC strength & $\phi$ & $\pi/50$ \\ \hline\hline
    Tunneling rate & $J_0$ & $\approx300\times2\pi$ Hz \\
    On-site interaction strength & $U$ & $\approx1.4\times2\pi$ kHz \\
    Collective interaction strength & $\xi$
    & $\approx14\times2\pi$ Hz \\
    Squeezing strength & $\chi$ & $\approx21\times2\pi$ mHz \\
    Optimal squeezing time & $t_{\t{opt}}^{\t{OAT}}$
    & $\approx0.66$ seconds \\ \hline
  \end{tabular}
\end{table}

@@

Rather than pulsing the clock laser, we can apply a continuous drive
to realize the Hamiltonian
\begin{align}
  H_{\t{drive}} = H_{\t{free}} + H_{\t{clock}}
  = -\xi\v S\c\v S - \bar h S_z - \Omega S_x - \sum_n b_n s_n^z,
\end{align}
where the drive strength $\Omega$ may generally depend on time.  As
shown in Appendix \ref{sec:effective_squeezing_drive}, the addition of
a drive with $\abs{\Omega}\ll N\abs{\xi}$ does not modify the
squeezing behavior of the spin-orbit coupling terms in
$H_{\t{drive}}$, whose effective action on the $S=N/2$ manifold is
captured by
\begin{align}
  H_{\t{drive,eff}} = -\bar h S_z - \Omega S_x + \chi S_z^2.
  \label{eq:H_drive_eff}
\end{align}
To follow the continuous-drive two-axis twisting prescription in
ref.~\cite{huang2015twoaxis}, we now
\begin{enumerate*}[label=(\roman*)]
\item blue-detune the clock laser by $\bar h$ and shift into an
  appropriate rotating frame to eliminate the $\sim S_z$ term in
  \eqref{eq:H_drive_eff},
\item modulate the drive as $\Omega\p{t}=\beta\omega\cos\p{\omega t}$
  with $\omega\gg N\chi$, and
\item move into the rotating frame of $-\Omega\p{t}S_x$.
\end{enumerate*}
After a secular approximation which relies on $\omega\gg N\chi$, this
procedure results in the effective Hamiltonian
\begin{align}
  H_{\t{drive,eff}}^{\t{mod}}
  = \f{\chi}{2} \p{\sp{\J_0\p{2\beta}+1} S_z^2
    - \sp{\J_0\p{2\beta}-1} S_y^2},
\end{align}
where $\J_0$ is the zero-order Bessel function of the first kind.
Choosing a modulation index $\beta$ such that $\J_0\p{2\beta}=1/3$ or
$\J_0\p{2\beta}=-1/3$ then respectively results in the two-axis
twisting Hamiltonians
\begin{align}
  H_{\t{TAT}}^{z,x}
  = \f{\chi}{3} \p{2 S_z^2 + S_y^2}
  \simeq \f{\chi}{3} \p{S_z^2 - S_x^2},
  &&
  H_{\t{TAT}}^{y,x}
  = \f{\chi}{3} \p{S_z^2 + 2 S_y^2}
  \simeq \f{\chi}{3}\p{S_y^2 - S_x^2},
  \label{eq:H_TAT_drive}
\end{align}
where $\simeq$ denotes equality up to a global energy shift. Given the
set of parameters in Table \ref{tab:parameters}, the two-axis twisting
Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with
$\omega=N\abs{\xi\chi}^{1/2}\approx27\times2\pi$ Hz and $\beta=0.906$.
The optimal squeezing time for this protocol is
$t_{\t{opt}}^{\t{TAT}}\approx0.98$ seconds.  A comparison of squeezing
through $H_{\t{OAT}}$ and $H_{\t{TAT}}^{z,x}$ via the parameters in
Table \ref{tab:parameters} is provided in Figure
\ref{fig:squeezing_comparison}.

\begin{figure}[h]
  \centering
  \includegraphics{squeezing_comparison.pdf}
  \caption{Comparison of squeezing through $H_{\t{OAT}}$ and
    $H_{\t{TAT}}^{z,x}$ via the parameters in Table
    \ref{tab:parameters} and the continuous two-axis twisting protocol
    with $\omega=27\times2\pi$ Hz and $\beta=0.906$.}
  \label{fig:squeezing_comparison}
\end{figure}
